---
title: "Home Data Science Challenge"
output: html_notebook
---

# Introduction

The goal of this challenge is to come up with the best price for a flat with these characteristics:

* Address: Almstadtstraße 9/11, 10119 Berlin
* Size: 78 sqm
* Floor: 1st
* Rooms: 2
* Built-in kitchen: yes
* Balcony: yes
* Construction year: 1902
* Condition: good
* Quality: good


# Exploratory Analysis and Data Wrangling

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(FNN)
```


```{r}
df = read.csv("HomeDataScienceChallenge.csv", sep = ";", dec = ",")
head(df)
```

The dataset contains information on Kaltmiete, Nebenkosten, Heizungkosten and the overall cost. The overall cost is missing for 3 homes. Given the very small dataset, every data point is precious therefore I will focus on predicting Kaltmiete instead of the overall cost. 

The Scikit Learn ML guide agrees with me :P. If your dataset < 50 samples -> Get more data!
http://scikit-learn.org/stable/tutorial/machine_learning_map/index.html.

```{r}
summary(df[c("kosten", "kstn_miete_gesamt")])
```

The flat for which a prediction needs to be made has only a few features available:

* address
* size
* floor
* rooms
* built in kitchen
* balcony 
* construction year
* quality
* condition

Therefore, the prediction model should only include these features. 

```{r}
df_mod = df[c(
  "oadr_strasse",
  "flaeche",
  "etage",
  "anz_zimmer",
  "aus_kueche_einbau_janein",
  "aus_balkon_terrasse_janein",
  "baujahr",
  "aus_klassen_empirica",
  "zust_klassen_empirica",
  "kosten"
  )]
```

## Preprocessing and Imputation
It's important to do some data preprocessing to make the model work. These features are boolean and shouldn't be considered as numeric ones:

* aus_kueche_einbau_janein
* aus_balkon_terrasse_janein

In R boolean features can be represented with the data type *factor*.
```{r}
df_mod$aus_kueche_einbau_janein = as.factor(df_mod$aus_kueche_einbau_janein)
df_mod$aus_balkon_terrasse_janein = as.factor(df_mod$aus_balkon_terrasse_janein)
summary(df_mod)
```

There are some missing features in the dataset: the building year is not available for 8 homes. The missing points will be imputed with the median of the building year.

```{r}
df_mod[is.na(df_mod$baujahr), "baujahr"] <- median(df_mod$baujahr, na.rm = T)
summary(df_mod)
```

## Feature Analysis

The plot shows that the data supports the relationship between the rent price and:

* number of rooms (1.5 rooms seems to be an outlier in the trend, but only because there's just 1 flat with 1.5 rooms)
* terrace/balcony (the average price being 114 Euro higher for a flat with a balcony)
* quality
* built in kitchen
* condition

For the floor of a building we observe a weird trend, probably because of the lack of data. It's hard to believe that the average rent price decreases for buildings on the third floor or higher.

```{r}
df_mod %>%
  gather(-oadr_strasse, -kosten, -flaeche, -baujahr, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y = kosten, col = flaeche)) + 
  geom_point() + facet_wrap(~ var, scales = "free") +
  theme(text = element_text(size=10), axis.text.x = element_text(angle=90, hjust=1)) + 
  scale_colour_gradient(low = "white",high = "red") +
  stat_summary(fun.y = mean, fun.ymin = mean, fun.ymax = mean, geom = "crossbar", width = 0.5)
```

The plot shows that the relationship between building year and rent price depends on the time period.

```{r}
df_mod[, c("kosten", "flaeche", "baujahr")] %>%
  gather(-kosten, key = "var", value = "value") %>%
  ggplot(aes(x = value, y = kosten)) + geom_point() + geom_smooth() +
  facet_wrap( ~ var, scales = "free") +
  theme(text = element_text(size = 10),
  axis.text.x = element_text(angle = 90, hjust = 1))
```

The relation between building year and rent price is best described by a picewise function.

```{r}
qplot(
  baujahr,
  kosten,
  group = baujahr < 1930,
  geom = c('point', 'smooth'),
  method = 'lm',
  se = F,
  data = df_mod
  )
```

Let's create a feature that captures this relation.

```{r}
df_mod$baujahr_below1930 <- ifelse(df_mod$baujahr <= 1930, 0, 1)
df_mod$baujahr_interaction <- df_mod$baujahr_below1930*(df_mod$baujahr-1930)
```

# Model Training
As a first try, I will build a linear regression model. 

## Validation Procedure
Splitting data into a training and test set is necessary in order to avoid overfitting (the model memorises instead of learns).

```{r}
set.seed(100)
train.ix <- sample(1:nrow(df_mod), size = 0.8*nrow(df_mod), replace = F)
df_train <- df_mod[train.ix,]
df_test <- df_mod[-train.ix,]
```

The train test price distribution shouldn't be too different, but it's difficult to ensure this with so few data points.

```{r}
par(mfrow = c(1, 3))
boxplot(df_train$kosten, ylab="Price", main="Training Data")
boxplot(df_test$kosten,  ylab="Price", main="Test Data")
boxplot(df$kosten,  ylab="Price", main="All Data")
```

## Linear Regression


```{r}
features <- c('flaeche', 'anz_zimmer', 'baujahr',  'baujahr_interaction',
              'aus_kueche_einbau_janein', 'aus_balkon_terrasse_janein', 
              'aus_klassen_empirica', 'zust_klassen_empirica')
features
```

```{r}
model_formula <- formula(df_mod[c('kosten', features)])
model_formula
```

This model leads to some strange results: e.g. number of rooms seems to have a negative effect on the price which is obviously wrong.
The problem is that there's a high correlation between the number of rooms and flat size (as expected). A model like linear regression may lead to meaningless results in such a situation. To deal with this, I will try a non parametric Machine Learning Model called Kth Nearest Neighbor Regression.

```{r}
lin_reg = lm(model_formula,
             data = df_train
)
summary(lin_reg)
```


```{r}
cor(df_mod$flaeche, df_mod$anz_zimmer)
```

### Model Validation

The plots show that there are some outliers in the data. Due to the very small amount of data I decided not to remove these outliers.

```{r}
plot(lin_reg)
```

### Error Analysis

For the homes where the predicted rent is higher than the actual one, the average error is 132 Euro.
The error is much higher for when we under-predict: on average being 209 Euro.
In terms of the percentage of the home price, the error is on average 17.8%.
We have to keep in mind that we only have 10 data points for analyzing the error :), so the statistics might not be very reliable.

```{r}
df_test$predicted_cost = predict(lin_reg, newdata=df_test)
df_test$error = df_test$predicted_cost - df_test$kosten
df_test$perc_error = abs(df_test$predicted_cost - df_test$kosten)/df_test$kosten
mean(df_test$perc_error)
```


```{r}
cor(df_test$kosten, df_test$predicted_cost)
```

```{r}
mean(df_test$error[df_test$error>0])
```

```{r}
mean(df_test$error[df_test$error<0])
```

```{r}
mean(abs(df_test$error))
```

### Prediction
The exciting part: what kind of rent will the model predict for our lovely new flat?
The prediction from linear regression is 1225 Euro!

```{r}
df_pred = data.frame(
  flaeche = 78,
  etage = 1,
  avg_room_size = 78/2,
  anz_zimmer = 2,
  aus_kueche_einbau_janein = 1,
  aus_balkon_terrasse_janein = 1,
  baujahr = 1902,
  baujahr_interaction = 0,
  aus_klassen_empirica = 'gut',
  zust_klassen_empirica = 'gut',
  stringsAsFactors = F
  )
df_pred$aus_kueche_einbau_janein <- as.factor(df_pred$aus_kueche_einbau_janein)
df_pred$aus_balkon_terrasse_janein <- as.factor(df_pred$aus_balkon_terrasse_janein)
df_pred
```

```{r}
df_pred$predicted_cost = predict(lin_reg, newdata=df_pred)
df_pred$predicted_cost
```

## Kth Nearest Neighbor Regression

We saw in the previous section that a statistical model like linear regression gave strange results in terms of the effect of the number of rooms on the flat price. To solve the present data issues I try out a Machine Learning model.

This leads to a much better prediction! The percentage error drops from 17.87% to 14.5%.
This model tends to have higher errors when we overestimate the rent price, while the linear model had lower errors in the case of rent price underestimation.

```{r}
x_train <- model.matrix( ~ .-1, df_train[features])
x_train <- subset(x_train, select = -aus_kueche_einbau_janein0)
x_test <- model.matrix( ~ .-1, df_test[features])
x_test <- subset(x_test, select = -aus_kueche_einbau_janein0)
x_full <- model.matrix( ~ .-1, df_mod[features])
x_full <- subset(x_full, select = -aus_kueche_einbau_janein0)

knn <- knn.reg(train = x_train, test = x_test, y = df_train$kosten, k = 3)
df_test$knn_predicted_cost <- knn$pred
df_test$knn_error = df_test$knn_predicted_cost - df_test$kosten
df_test$knn_perc_error = abs(df_test$knn_predicted_cost - df_test$kosten)/df_test$kosten
mean(df_test$knn_perc_error)
```

```{r}
mean(df_test$knn_error[df_test$knn_error>0])
```

```{r}
mean(df_test$knn_error[df_test$knn_error<0])
```

```{r}
mean(abs(df_test$knn_error))
```



## Prediction

KNN Regression predicts a price of **1145 Euro** for our lovely new flat.

```{r}
x_pred = data.frame(
  flaeche = 78,
  anz_zimmer = 2,
  aus_kueche_einbau_janein1 = 1,
  aus_balkon_terrasse_janein1 = 1,
  baujahr = 1902,
  baujahr_interaction = 0,
  aus_klassen_empiricagut = 1,
  aus_klassen_empiricahochwertig = 0,
  aus_klassen_empiricanormal = 0,
  zust_klassen_empiricanormal = 0,
  zust_klassen_empiricaschlecht = 0
  )
x_pred = as.matrix(x_pred[colnames(x_train)])
knn.reg(train = x_full, test = x_pred, y = df_mod$kosten, k = 3)

```
